Errata

### 2025-04-05 tex/Manuscript-2.tex
Update Dissertation and manuscript files with new content and formatting improvements

In addition, under humid conditions, biocrusts have a higher proportion of fungi and bryophytes, resulting in rough surfaces with a remarkable capacity to store capillary water, increased pore space in their structure and enhanced carbon fixation \citer{RiverasMunoz2022,Weber2022}. These changes in surface roughness affect the surface hydrology of the soil by modifying the residence time of water on the surface \citep{Kidron2022}. Changes in water residence time, in turn, affect the distribution of infiltration and runoff, alter surface redox conditions, modify soil erodibility, and regulate the release of solutes into water \citep{Kalnejais2010}. This emphasizes the ability of biocrusts to modify water and sediment fluxes. In particular, lichen-dominated dryland biocrusts can enhance surface runoff and reduce sediment yield by surface clogging and subsequently surface saturation, and runoff initiation \citep{Kidron2021}. They can also dramatically alter albedo, soil temperature dynamics and thus consequently the potential evapotranspiration \citep{Liu2020b,Rutherford2017,Whitney2017,Xiao2019}. Moss-dominated biocrusts with a rough morphology have a high water-holding potential and can decrease both runoff and sediment yield \citep{Juan2023,Seitz2017,Silva2019,Zeng2025}. Other effects of biocrusts that also include near-surface infiltration within the biocrust cover and its effect on percolation report enhanced infiltration and percolation in semi-arid ecosystems \citep{Chamizo2016} and reclaimed soils \citep{Gypser2016}, alteration rainwater flow in drylands \citep{Li2022}, and reduced infiltration while increasing erosion resistance in disturbed forests \citep{Szyja2023}. Moreover, comparative experimental studies focusing on soil water movement through different types of biocrust remain limited.

@@

Infiltration boxes (Figure \ref{fig:M2-F3}) were developed as part of a rainfall simulation experiment to measure runoff and percolation flow, including their matter content. Undisturbed soil samples were collected using cutting frames (\SI{20}{\centi\metre} \times \SI{30}{\centi\metre} \times \SI{7}{\centi\metre}, Figure \ref{fig:M2-F3}a), carefully installed with minimal surface and subsurface disturbances (Figure \ref{fig:M2-F3}c, Seitz (2015)). Cutting frames are made from \SI{1}{\milli\meter} metal plates, sharpened in the bottom sides, and include an upper border to transfer the strength to them without direct contact with the soil and a lower one to stop the frame from burying beyond the desired height of the sample (Figure \ref{fig:M2-F3}a). Subsequently, the cutting frames were excavated around, and a metal plate was inserted underneath it (Figure \ref{fig:M2-F3}d). The cutting frames with the soil samples were covered with metal plates, wrapped in plastic foil, and carefully transported to a flat area with water available. Then, the wrappings were removed, and the frames with the soil samples were stacked on a permeable metal plate and placed inside the soil erosion flux box (Figure \ref{fig:M2-F3}a) designed as steel containers with a triangular surface runoff gutter and an outlet at the bottom to capture the percolation flow (Figure \ref{fig:M2-F3}a). Soil depth within the boxes is \SI{7}{\centi\meter}. In the case of the presence of small plants, they were cut flush with the surface using a scissor and paying attention to not pull them and avoid surface disturbances. The water content of each sample was measured by a TDR probe (Delta-T Devices Ltd. Cambridge, UK) when sampling, using the average of 3 measurements directly next to the sampling area. Perpendicular photographs were taken on each sample with a digital camera (Sony ILCE-6000 equipped with a lens SELP1650, Tokyo, Japan) and processed with the grid quadrat method overlying a digital grid of 100 subdivisions and separating biocrusts by visual inspection \citep{Belnap2001} to assess the biocrust ground cover.

@@

BSC affected dissolved C and N fractions (DOC, DON) across the climatic gradient. Site-specific conditions and biocrust interactions strongly influenced DOC in runoff, which explained 56.0\% of the variability ($p = 1.939e^{-15}$). Biocrusts amplified DOC in some areas, such as in SG, while reducing it in others, such as NA. Similar patterns were observed in percolation, where site and biocrust interactions accounted for 33.1\% of the variability ($p < 1.173e^{−8}$). DON in runoff showed analogous trends, with significant site-biocrust interactions explaining 56.3\% of the variability ($p = 2.2e^{−16}$). For instance, in northern sites like SG, biocrusts increased DON from \SI[separate-uncertainty,multi-part-units=single]{1.3 \pm 1.0}{ppm} to \SI[separate-uncertainty,multi-part-units=single]{2.2 \pm 2.0}{ppm}. However, in the southest site NA, biocrust decreased DON from \SI[separate-uncertainty,multi-part-units=single]{0.6 \pm 0.7} to \SI[separate-uncertainty,multi-part-units=single]{0.3 \pm 0.8}{ppm}. In percolation, DON content followed similar trends, with 45.6\% of variability explained by site and biocrust interactions ($p = 3.905e^{−13}$).

>>>>

In addition, under humid conditions, biocrusts have a higher proportion of fungi and bryophytes, resulting in rough surfaces with a remarkable capacity to store capillary water, increased pore space in their structure and enhanced carbon fixation \citep{RiverasMunoz2022,Weber2022}. These changes in surface roughness affect the surface hydrology of the soil by modifying the residence time of water on the surface \citep{Kidron2022}. Changes in water residence time, in turn, affect the distribution of infiltration and runoff, alter surface redox conditions, modify soil erodibility, and regulate the release of solutes into water \citep{Kalnejais2010}. This emphasizes the ability of biocrusts to modify water and sediment fluxes. In particular, lichen-dominated dryland biocrusts can enhance surface runoff and reduce sediment yield by surface clogging and subsequently surface saturation, and runoff initiation \citep{Kidron2021}. They can also dramatically alter albedo, soil temperature dynamics and thus consequently the potential evapotranspiration \citep{Liu2020b,Rutherford2017,Whitney2017,Xiao2019}. Moss-dominated biocrusts with a rough morphology have a high water-holding potential and can decrease both runoff and sediment yield \citep{Juan2023,Seitz2017,Silva2019,Zeng2025}. Other effects of biocrusts that also include near-surface infiltration within the biocrust cover and its effect on percolation report enhanced infiltration and percolation in semi-arid ecosystems \citep{Chamizo2016} and reclaimed soils \citep{Gypser2016}, alteration rainwater flow in drylands \citep{Li2022}, and reduced infiltration while increasing erosion resistance in disturbed forests \citep{Szyja2023}. Moreover, comparative experimental studies focusing on soil water movement through different types of biocrust remain limited.

@@

Infiltration boxes (Figure \ref{fig:M2-F3}) were developed as part of a rainfall simulation experiment to measure runoff and percolation flow, including their matter content. Undisturbed soil samples were collected using cutting frames (\SI{20}{\centi\metre} $\times$ \SI{30}{\centi\metre} $\times$ \SI{7}{\centi\metre}, Figure \ref{fig:M2-F3}a), carefully installed with minimal surface and subsurface disturbances (Figure \ref{fig:M2-F3}c, Seitz (2015)). Cutting frames are made from \SI{1}{\milli\meter} metal plates, sharpened in the bottom sides, and include an upper border to transfer the strength to them without direct contact with the soil and a lower one to stop the frame from burying beyond the desired height of the sample (Figure \ref{fig:M2-F3}a). Subsequently, the cutting frames were excavated around, and a metal plate was inserted underneath it (Figure \ref{fig:M2-F3}d). The cutting frames with the soil samples were covered with metal plates, wrapped in plastic foil, and carefully transported to a flat area with water available. Then, the wrappings were removed, and the frames with the soil samples were stacked on a permeable metal plate and placed inside the soil erosion flux box (Figure \ref{fig:M2-F3}a) designed as steel containers with a triangular surface runoff gutter and an outlet at the bottom to capture the percolation flow (Figure \ref{fig:M2-F3}a). Soil depth within the boxes is \SI{7}{\centi\meter}. In the case of the presence of small plants, they were cut flush with the surface using a scissor and paying attention to not pull them and avoid surface disturbances. The water content of each sample was measured by a TDR probe (Delta-T Devices Ltd. Cambridge, UK) when sampling, using the average of 3 measurements directly next to the sampling area. Perpendicular photographs were taken on each sample with a digital camera (Sony ILCE-6000 equipped with a lens SELP1650, Tokyo, Japan) and processed with the grid quadrat method overlying a digital grid of 100 subdivisions and separating biocrusts by visual inspection \citep{Belnap2001} to assess the biocrust ground cover.

@@

BSC affected dissolved C and N fractions (DOC, DON) across the climatic gradient. Site-specific conditions and biocrust interactions strongly influenced DOC in runoff, which explained 56.0\% of the variability ($p = 1.939e^{-15}$). Biocrusts amplified DOC in some areas, such as in SG, while reducing it in others, such as NA. Similar patterns were observed in percolation, where site and biocrust interactions accounted for 33.1\% of the variability ($p < 1.173e^{-8}$). DON in runoff showed analogous trends, with significant site-biocrust interactions explaining 56.3\% of the variability ($p = 2.2e^{-16}$). For instance, in northern sites like SG, biocrusts increased DON from \SI[separate-uncertainty,multi-part-units=single]{1.3 \pm 1.0}{ppm} to \SI[separate-uncertainty,multi-part-units=single]{2.2 \pm 2.0}{ppm}. However, in the southest site NA, biocrust decreased DON from \SI[separate-uncertainty,multi-part-units=single]{0.6 \pm 0.7} to \SI[separate-uncertainty,multi-part-units=single]{0.3 \pm 0.8}{ppm}. In percolation, DON content followed similar trends, with 45.6\% of variability explained by site and biocrust interactions ($p = 3.905e^{-13}$).
